---
title: "4_report"
author: "Giang Pham"
date: "2023-05-06"
output: html_document
---

BEFORE:
      1_data_process.R
      2_data_explore.Rmd
      3_data_analyze.R

AFTER:
      report folder
      4_report.html
      
# 1. SETUP
```{r message=FALSE, warning=FALSE}
# load packages
library(tidyverse)   # process dataframe
library(effects)     # present effects
library(patchwork)   # arrange plot
library(broom.mixed) # summarize estimation
library(RColorBrewer)# visualize
library(writexl)     # write excel
library(reactable)   # interactive table
    

# load function
## centering function
#c. <- function (x) {(x - mean(x, na.rm = T))} 
## log, centering function
#clog. <- function (x) {(log(x) - mean(log(x), na.rm = T))} 

# load data
dir_data <- "./data"
dir_model <- "./model"
dir_report <- "./report"
```

# 2. MODEL RESULTS

## data
```{r message=FALSE, warning=FALSE}
# data
sdg381 <- read_csv(file.path(dir_data, "@data_sdg381.csv"))
sdg381_rmnch <- read_csv(file.path(dir_data, "@data_sdg381_rmnch.csv"))
sdg382 <- read_csv(file.path(dir_data, "@data_sdg382.csv"))

# transform data
## note: prepaid.private.health.spending has 0 -> cannot log -> for 0 value, add 0.0001 
## as the smallest value of total.aid.for.health 

## sdg381
sdg381 <- sdg381 %>% 
  filter(year < 2019) %>%
  mutate(#response variable
         log_sdg381 = log(sdg381),
         # explanatory variable
         log_total.aid = log(total.aid), 
         log_total.aid.volatility = log(total.aid.volatility),
         log_total.aid.for.health = log(total.aid.for.health), 
         log_total.aid.for.health.volatility = log(total.aid.for.health.volatility),
         log_total.aid.for.health_rmnch = log(total.aid.for.health_rmnch), 
         log_total.aid.for.health.volatility_rmnch = log(total.aid.for.health.volatility_rmnch),
         institutional.index = institutional.index,
         # control variable
         log_gdp.per.capita = log(gdp.per.capita), 
         log_government.health.spending = log(government.health.spending), 
         log_prepaid.private.health.spending = if_else(prepaid.private.health.spending == 0,
                                                       log(prepaid.private.health.spending + 0.0001),
                                                       log(prepaid.private.health.spending)),
         log_out.of.pocket.health.spending = log(out.of.pocket.health.spending), 
         #log_health.worker = log(health.worker), 
         log_sdi = log(sdi),
         # random variables
         fyear = factor(year))

## sdg381_rmnch
sdg381_rmnch <- sdg381_rmnch %>% 
  #filter(year < 2019) %>%
  mutate(#response variable
    log_sdg381_rmnch = log(sdg381_rmnch),
    # explanatory variable
    log_total.aid = log(total.aid), 
    log_total.aid.volatility = log(total.aid.volatility),
    log_total.aid.for.health = log(total.aid.for.health), 
    log_total.aid.for.health.volatility = log(total.aid.for.health.volatility),
    log_total.aid.for.health_rmnch = log(total.aid.for.health_rmnch), 
    log_total.aid.for.health.volatility_rmnch = log(total.aid.for.health.volatility_rmnch),
    institutional.index = institutional.index,
    # control variable
    log_gdp.per.capita = log(gdp.per.capita), 
    log_government.health.spending = log(government.health.spending), 
    log_prepaid.private.health.spending = if_else(prepaid.private.health.spending == 0,
                                                  log(prepaid.private.health.spending + 0.0001),
                                                  log(prepaid.private.health.spending)),
    log_out.of.pocket.health.spending = log(out.of.pocket.health.spending), 
    #log_health.worker = log(health.worker), 
    log_sdi = log(sdi),
    # random variables
    fyear = factor(year))

## sdg382
sdg382 <- sdg382 %>% 
  #filter(year < 2019) %>%
  mutate(#response variable
    log_sdg382 = log(sdg382),
    # explanatory variable
    log_total.aid = log(total.aid), 
    log_total.aid.volatility = log(total.aid.volatility),
    log_total.aid.for.health = log(total.aid.for.health), 
    log_total.aid.for.health.volatility = log(total.aid.for.health.volatility),
    log_total.aid.for.health_rmnch = log(total.aid.for.health_rmnch), 
    log_total.aid.for.health.volatility_rmnch = log(total.aid.for.health.volatility_rmnch),
    institutional.index = institutional.index,
    # control variable
    log_gdp.per.capita = log(gdp.per.capita), 
    log_government.health.spending = log(government.health.spending), 
    log_prepaid.private.health.spending = if_else(prepaid.private.health.spending == 0,
                                                  log(prepaid.private.health.spending + 0.0001),
                                                  log(prepaid.private.health.spending)),
    log_out.of.pocket.health.spending = log(out.of.pocket.health.spending), 
    #log_health.worker = log(health.worker), 
    log_sdi = log(sdi),
    # random variables
    fyear = factor(year))

``` 

## Plot lme results

### sdg382

```{r}
# model result
sdg382_base <- read_rds(file.path(dir_model, "sdg382_base.rds"))
sdg382_dah <- read_rds(file.path(dir_model, "sdg382_total.aid.for.health.rds"))
sdg382_dah.vol <- read_rds(file.path(dir_model, "sdg382_total.aid.for.health.vol.rds"))
```

#### baseline 

##### setup

```{r}
# data
data_base <- sdg382 %>% filter(is.na(log_sdg382) == F,
                               is.na(log_gdp.per.capita) == F, 
                               is.na(log_government.health.spending) == F, 
                               is.na(log_prepaid.private.health.spending) == F,
                               is.na(log_out.of.pocket.health.spending) == F, 
                               #is.na(log_health.worker) == F, 
                               is.na(log_sdi) == F)
data <- data_base

# model summary to select variables to plot
summary(sdg382_base)

# create empty plot list
plot_list <- list()
```

##### log_gdp.per.capita  
```{r}
# setup
pred <- as.data.frame(Effect(c("log_gdp.per.capita"), 
                      sdg382_base, 
                      xlevels = 20)) 
pred <- pred %>%
  mutate(var = exp(pred[,1]),
         fit_trans = exp(fit),
         lower_trans = exp(lower),
         upper_trans = exp(upper))

# plot
ggplot() + 
  geom_line(data = pred, aes(x = var, 
                             y = fit_trans )) +
  geom_ribbon(data = pred, aes(x = var,
                               ymin = lower_trans,
                              ymax = upper_trans),
              alpha = 0.3) +
  theme_classic() +
  labs(x = "GDP per capita (2015 USD)",
       y = "SDG 3.8.2")

plot_list <- append(plot_list, list(last_plot()))

# summary
tibble(fit_at.min = mean(filter(pred, var == min(var))$fit),
       fit_at.max = mean(filter(pred, var == max(var))$fit)) %>% 
  mutate(fit_at.min = exp(fit_at.min),
       fit_at.max = exp(fit_at.max),
       fit_diff.percent = (fit_at.max - fit_at.min)/fit_at.min*100) %>%
  round(2)

pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  mutate(fit_trans = round(fit_trans, 2),
         lower_trans = round(lower_trans, 2),
         upper_trans = round(upper_trans, 2)) %>%
  mutate(CI95 = paste0(lower_trans, "-", upper_trans)) %>%
  select(range, fit_trans, CI95) 

```
##### log_government.health.spending
```{r}
# setup
pred <- as.data.frame(Effect(c("log_government.health.spending"), 
                      sdg382_base, 
                      xlevels = 20)) 
pred <- pred %>%
  mutate(var = exp(pred[,1]),
         fit_trans = exp(fit),
         lower_trans = exp(lower),
         upper_trans = exp(upper))

# plot
ggplot() + 
  geom_line(data = pred, aes(x = var, 
                             y = fit_trans )) +
  geom_ribbon(data = pred, aes(x = var,
                               ymin = lower_trans,
                              ymax = upper_trans),
              alpha = 0.3) +
  theme_classic() +
  labs(x = "Government Health Spending per GDP (%)",
       y = "SDG 3.8.2")

plot_list <- append(plot_list, list(last_plot()))

# summary
tibble(fit_at.min = mean(filter(pred, var == min(var))$fit),
       fit_at.max = mean(filter(pred, var == max(var))$fit)) %>% 
  mutate(fit_at.min = exp(fit_at.min),
       fit_at.max = exp(fit_at.max),
       fit_diff.percent = (fit_at.max - fit_at.min)/fit_at.min*100) %>%
  round(2)

pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  mutate(fit_trans = round(fit_trans, 2),
         lower_trans = round(lower_trans, 2),
         upper_trans = round(upper_trans, 2)) %>%
  mutate(CI95 = paste0(lower_trans, "-", upper_trans)) %>%
  select(range, fit_trans, CI95) 

```
##### log_out.of.pocket.health.spending
```{r}
# setup
pred <- as.data.frame(Effect(c("log_out.of.pocket.health.spending"), 
                      sdg382_base, 
                      xlevels = 20)) 
pred <- pred %>%
  mutate(var = exp(pred[,1]),
         fit_trans = exp(fit),
         lower_trans = exp(lower),
         upper_trans = exp(upper))

# plot
ggplot() + 
  geom_line(data = pred, aes(x = var, 
                             y = fit_trans )) +
  geom_ribbon(data = pred, aes(x = var,
                               ymin = lower_trans,
                              ymax = upper_trans),
              alpha = 0.3) +
  theme_classic() +
  labs(x = "Out-of-pocket Health Spending per GDP (%)",
       y = "SDG 3.8.2")

plot_list <- append(plot_list, list(last_plot()))

# summary
tibble(fit_at.min = mean(filter(pred, var == min(var))$fit),
       fit_at.max = mean(filter(pred, var == max(var))$fit)) %>% 
  mutate(fit_at.min = exp(fit_at.min),
       fit_at.max = exp(fit_at.max),
       fit_diff.percent = (fit_at.max - fit_at.min)/fit_at.min*100) %>%
  round(2)

pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  mutate(fit_trans = round(fit_trans, 2),
         lower_trans = round(lower_trans, 2),
         upper_trans = round(upper_trans, 2)) %>%
  mutate(CI95 = paste0(lower_trans, "-", upper_trans)) %>%
  select(range, fit_trans, CI95) 
```
##### save plot

```{r}
p_patch <- ( plot_list[[1]] | plot_list[[2]] | plot_list[[3]]) + plot_annotation(tag_levels = 'A') & ylab(NULL) 

p_patch <- wrap_elements(panel = p_patch) +
  labs(tag = "SDG 3.8.2") +
  theme(plot.tag = element_text(size = rel(1), angle = 90),
        plot.tag.position = "left")

p_patch

# save plot
ggsave(filename = "./report/sdg382_baseline.png", 
       plot = p_patch,
       width = 28,
       height = 10,
       units = "cm",
       dpi = 600)
```

####  total aid for health ---------------
##### setup
```{r}
# data
data_total.aid.for.health <- sdg382 %>% filter(is.na(log_sdg382) == F,
                                               is.na(log_total.aid.for.health) == F,
                                               is.na(institutional.index) == F,
                                               is.na(log_gdp.per.capita) == F, 
                                               is.na(log_government.health.spending) == F, 
                                               is.na(log_prepaid.private.health.spending) == F,
                                               is.na(log_out.of.pocket.health.spending) == F, 
                                               #is.na(log_health.worker) == F, 
                                               is.na(log_sdi) == F)

data <- data_total.aid.for.health

# model summary to select variables to plot
summary(sdg382_dah)

# create empty plot list
plot_list <- list()
```
##### log_total.aid.for.health:institutional.index 
```{r}
# setup
pred <- as.data.frame(Effect(c("log_total.aid.for.health", "institutional.index"), 
                      sdg382_dah,
                      xlevels = 21))

df_inst <- tibble(level = factor(c("low", "middle", "high"), levels = c("low", "middle", "high")),
                 institutional.index = unique(pred$institutional.index)[c(1,11,21)])

pred <- pred %>%
  mutate(var = exp(pred[,1]),
         fit_trans = exp(fit),
         lower_trans = exp(lower),
         upper_trans = exp(upper)) %>%
  filter(institutional.index %in% df_inst$institutional.index) %>%
  left_join(df_inst)

# plot
ggplot() + 
  geom_line(data = pred, aes(x = var, 
                             y = fit_trans,
                             )) +
  geom_ribbon(data = pred, aes(x = var,
                               ymin = lower_trans,
                              ymax = upper_trans),
              alpha = 0.3) +
  theme_classic() +
  labs(x = "DAH per GDP (%)",
       y = "SDG 3.8.2") +
  facet_wrap(~ level)

plot_list <- append(plot_list, list(last_plot())) 

# summary
pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  select(fit_trans, range, level) %>%
  pivot_wider(names_from = range, values_from = fit_trans) %>%
  mutate(fit_diff.percent = (fit_at.max - fit_at.min)/fit_at.min*100) %>% 
  mutate(fit_at.min = round(fit_at.min, 2),
         fit_at.max = round(fit_at.max, 2),
         fit_diff.percent = round(fit_diff.percent, 2))

pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  mutate(fit_trans = round(fit_trans, 2),
         lower_trans = round(lower_trans, 2),
         upper_trans = round(upper_trans, 2)) %>%
  mutate(CI95 = paste0(lower_trans, "-", upper_trans)) %>%
  select(level, range, fit_trans, CI95) 

```

##### institutional.index  

```{r}
# setup
pred <- as.data.frame(Effect(c("institutional.index"), 
                      sdg382_dah, 
                      xlevels = 20)) 

pred <- pred %>%
  mutate(var = exp(pred[,1]),
         fit_trans = exp(fit),
         lower_trans = exp(lower),
         upper_trans = exp(upper))

# plot
ggplot() + 
  geom_line(data = pred, aes(x = var, 
                             y = fit_trans )) +
  geom_ribbon(data = pred, aes(x = var,
                               ymin = lower_trans,
                              ymax = upper_trans),
              alpha = 0.3) +
  theme_classic() +
  labs(x = "Institutional quality",
       y = "SDG 3.8.2")

plot_list <- append(plot_list, list(last_plot()))

# summary
tibble(fit_at.min = mean(filter(pred, var == min(var))$fit),
       fit_at.max = mean(filter(pred, var == max(var))$fit)) %>% 
  mutate(fit_at.min = exp(fit_at.min),
       fit_at.max = exp(fit_at.max),
       fit_diff.percent = (fit_at.max - fit_at.min)/fit_at.min*100) %>%
  round(2)

pred %>% 
  filter(var %in% c(min(var), max(var))) %>%
  mutate(range = if_else(var == min(var), "fit_at.min", "fit_at.max")) %>%
  mutate(fit_trans = round(fit_trans, 2),
         lower_trans = round(lower_trans, 2),
         upper_trans = round(upper_trans, 2)) %>%
  mutate(CI95 = paste0(lower_trans, "-", upper_trans)) %>%
  select(range, fit_trans, CI95) 
```

##### save plot

```{r}
( plot_list[[1]] )
# save plot
ggsave(filename = "./report/sdg382_dah_dah_inst.png", 
       plot = last_plot(),
       width = 18,
       height = 10,
       units = "cm",
       dpi = 600)

( plot_list[[2]] )
# save plot
ggsave(filename = "./report/sdg382_dah_inst.png", 
       plot = last_plot(),
       width = 18,
       height = 10,
       units = "cm",
       dpi = 600)
```

####  total aid for health volatility ---------------
##### setup
```{r}
# data
data_total.aid.for.health.vol <- sdg382 %>% filter(is.na(log_sdg382) == F,
                                                   is.na(log_total.aid.for.health.volatility) == F,
                                                   is.na(log_total.aid.for.health) == F,
                                                   is.na(institutional.index) == F,
                                                   is.na(log_gdp.per.capita) == F, 
                                                   is.na(log_government.health.spending) == F, 
                                                   is.na(log_prepaid.private.health.spending) == F,
                                                   is.na(log_out.of.pocket.health.spending) == F, 
                                                   #is.na(log_health.worker) == F, 
                                                   is.na(log_sdi) == F)

data <- data_total.aid.for.health.vol

# model summary to select variables to plot
summary(sdg382_dah.vol)

# create empty plot list
plot_list <- list()
```
