---
title: "4_report"
author: "Giang Pham"
date: "2023-05-06"
output: html_document
---

BEFORE:
      1_data_process.R
      2_data_explore.Rmd
      3_data_analyze.R

AFTER:
      report folder
      4_report.html
      
# 1. SETUP
```{r message=FALSE, warning=FALSE}
# load packages
library(tidyverse)   # process dataframe
library(effects)     # present effects
library(patchwork)   # arrange plot
library(broom.mixed) # summarize estimation
library(RColorBrewer)# visualize
library(writexl)     # write excel
library(reactable)   # interactive table
    

# load function
## centering function
#c. <- function (x) {(x - mean(x, na.rm = T))} 
## log, centering function
#clog. <- function (x) {(log(x) - mean(log(x), na.rm = T))} 

# load data
dir_data <- "./data"
dir_model <- "./model"
dir_report <- "./report"
```

# 2. DESCRIPTIVE RESULTS

load data
```{r message=FALSE, warning=FALSE}
sdg381 <- read_csv(file.path(dir_data, "@data_sdg381.csv")) %>% filter(year <= 2017)
sdg381_rmnch <- read_csv(file.path(dir_data, "@data_sdg381_rmnch.csv"))
sdg382 <- read_csv(file.path(dir_data, "@data_sdg382.csv")) 
data_explanatory <- read_csv(file.path(dir_data, "@data_explanatory.csv"))
```


## 2.1. descriptive results all variables
```{r}
# note: summarize mean, sd all vars

# data
## response vars
sum_res <- tibble(variable = c("sdg381", "sdg381_rmnch", "sdg382"),
       mean = c(mean(sdg381$sdg381, na.rm = T),
                mean(sdg381_rmnch$sdg381_rmnch, na.rm = T),
                mean(sdg382$sdg382, na.rm = T)),
       sd = c(sd(sdg381$sdg381, na.rm = T),
              sd(sdg381_rmnch$sdg381_rmnch, na.rm = T),
              sd(sdg382$sdg382, na.rm = T))) %>% 
  mutate(mean = round(mean, 2),
         sd = round(sd, 2))

## explanatory vars
sum_exp <- select(sdg381, 
                  health.worker:total.aid.for.health,
                 total.aid.for.health.volatility:institutional.index)
sum_exp <- tibble(variable = names(sum_exp), 
                  mean = apply(sum_exp, 2, mean, na.rm = T),
                  sd = apply(sum_exp, 2, sd, na.rm = T)) %>% 
  mutate(mean = round(mean, 2),
         sd = round(sd, 2))

## merge descriptive results
descriptive_results <- bind_rows(sum_res, sum_exp) 
write_xlsx(descriptive_results, file.path(dir_report, "descriptive_results.xlsx"))
```

```{r}
# table
reactable(descriptive_results, filterable = TRUE)
```

## 2.2. trend sdg381, sdg381_rmnch, sdg382
### sdg381

```{r}
# note: plot the change of value over time

# data 
## workflow
## 1. group_by country, income.group, region
## 2. summarize range year, value
## 3. mutate value diff, max
## 4. round values

trend_sdg381 <- sdg381 %>% 
  group_by(country, income.group_wb, region_un) %>%
  summarize(range_year = range(year),
            range_value = range(sdg381)) %>%
  mutate(value_diff = diff(range_value),
         value_max = max(range_value)) %>%
  mutate(range_value = round(range_value, 2),
         value_diff = round(value_diff, 2),
         value_max = round(value_max, 2))
```

```{r}
# table
reactable(trend_sdg381, filterable = TRUE)

# plot value + diff 
## setup
quantile(trend_sdg381$value_diff, probs = seq(0, 1, 0.1)) #check quantile
break_diff <- c(0, 10, 27, 32) #quantile 0, 10, 90, 100

ggplot(data = trend_sdg381, 
       aes(x = reorder(country, -value_max), 
           y = range_value)) + 
  geom_line(aes(color = cut(value_diff, breaks = break_diff)),
            linewidth = 1) +
  geom_point(aes(fill = factor(range_year)), shape = 21) +
  scale_color_brewer(palette = "BrBG") + #"Spectral"  
  scale_fill_brewer(palette = "BrBG", direction = -1) + #PuRd
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.1",
       color = "Index difference",
       fill = "Year")

# save plot
ggsave(filename = "./report/desc_sdg381_trend.png", 
       plot = last_plot(),
       width = 18,
       height = 24,
       units = "cm",
       dpi = 600)
````

```{r}
# plot diff
ggplot(data = trend_sdg381, 
       aes(x = reorder(country, -value_diff), 
           y = value_diff)) + 
  geom_point() +
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.1 - Index difference")
```

### sdg381_rmnch
```{r}
# note: plot the difference between poorest and richest group in the most recent year

# data
## workflow
## 1. filter non-NA values
## 2. filter only poorest and richest groups
## 3. group_by 
## 4. summarize year, value
## 5. mutate value diff, max
## 6. round values
## 7. mutate country.year - showing year with available data

trend_sdg381_rmnch <- sdg381_rmnch %>% 
  filter(is.na(sdg381_rmnch) == F) %>%
  filter(economic.group %in% c("Quintile 1 (poorest)", "Quintile 5 (richest)")) %>%
  group_by(country, income.group_wb, region_un, economic.group) %>%
  summarize(year = max(year),
            value = max(sdg381_rmnch)) %>%
  mutate(value_diff = diff(value),
         value_max = max(value)) %>%
  mutate(value = round(value, 2),
         value_diff = round(value_diff, 2),
         value_max = round(value_max, 2)) %>%
  mutate(country.year = paste0(country, " ", year))

```

```{r}
# table
reactable(trend_sdg381_rmnch, filterable = TRUE)

# plot value + diff 
## setup
quantile(trend_sdg381_rmnch$value_diff, probs = seq(0, 1, 0.1)) #check quantile
break_diff <- c(0, 10, 35, 50) #quantile 10, 90, 100

ggplot(data = trend_sdg381_rmnch, 
       aes(x = reorder(country.year, -value_max), 
           y = value)) + 
  geom_line(aes(color = cut(value_diff, breaks = break_diff)),
            linewidth = 1) +
  geom_point(aes(fill = economic.group), shape = 21) +
  scale_color_brewer(palette = "BrBG") + #"Spectral"
  scale_fill_brewer(palette = "BrBG", direction = -1) + #PuRd
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.1 RMNCH",
       color = "Index difference",
       fill = "Economic group")

# save plot
ggsave(filename = "./report/desc_sdg381_rmnch_trend.png", 
       plot = last_plot(),
       width = 18,
       height = 24,
       units = "cm",
       dpi = 600)

```

```{r}
# plot diff
ggplot(data = trend_sdg381_rmnch, 
       aes(x = reorder(country, -value_diff), 
           y = value_diff)) + 
  geom_point() +
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.1 RMNCH - Index difference")

```

### sdg382

```{r}
# note: plot the change of value over time

# data 
## workflow
## 1. group_by country, income.group, region
## 2. summarize range year, value
## 3. mutate range_year_cat because of different min, max year
## 4. mutate country.year - showing range_year
## 5. mutate value diff, max
## 6. round values
## 7. remove data without diff

trend_sdg382 <- sdg382 %>% 
  group_by(country, income.group_wb, region_un) %>%
  filter(year %in% c(min(year), max(year))) %>%
  summarize(range_year = range(year),
            range_value = range(sdg382)) %>%
  mutate(range_year_cat = if_else(range_year == min(range_year), "Year (first)", "Year (last)")) %>%
  mutate(country.year = paste0(country, " ", min(range_year), "-", max(range_year))) %>%
  mutate(value_diff = diff(range_value),
         value_max = max(range_value)) %>%
  mutate(range_value = round(range_value, 2),
         value_diff = round(value_diff, 2),
         value_max = round(value_max, 2)) %>%
  filter(value_diff != 0)
```

```{r}
# table
reactable(trend_sdg382, filterable = TRUE)

# plot value + diff 
## setup
quantile(trend_sdg382$value_diff, probs = seq(0, 1, 0.1)) #check quantile
break_diff <- c(0, 1, 10, 25) #quantile 0, 10, 90, 100

ggplot(data = trend_sdg382, 
       aes(x = reorder(country.year, -value_max), 
           y = range_value)) + 
  geom_line(aes(color = cut(value_diff, breaks = break_diff)),
            linewidth = 1) +
  geom_point(aes(fill = range_year_cat), shape = 21) +
  scale_color_brewer(palette = "BrBG") + #"Spectral"  
  scale_fill_brewer(palette = "BrBG", direction = -1) + #PuRd
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.2",
       color = "Index difference",
       fill = "Year") 

# save plot
ggsave(filename = "./report/desc_sdg382_trend.png", 
       plot = last_plot(),
       width = 18,
       height = 24,
       units = "cm",
       dpi = 600)
```

```{r}
# plot diff
ggplot(data = trend_sdg382, 
       aes(x = reorder(country.year, -value_diff), 
           y = value_diff)) + 
  geom_point() +
  coord_flip() +
  theme_classic() +
  labs(x = NULL,
       y = "SDG 3.8.2 - Index difference") 
```

## 2.3. health spending composition per total health spending

```{r}
# note: plot health spending composition per total health spending

# data 
## workflow
## 1. filter year 1995, 2017
## 2. select country, year, the vars
## 3. filter non NA values
## 4. rename
## 5. pivot longer

the <- data_explanatory %>% 
  filter(year %in% c(1995, 2019)) %>%
  select(country, year, gdp.per.capita, gdp, ghes_per_the:dah_per_the) %>%
  filter(is.na(gdp.per.capita) == F) %>%
  rename(`Government` = ghes_per_the,
         `Prepaid private` = ppp_per_the,
         `Out-of-pocket` = oop_per_the,
         `DAH` = dah_per_the) %>%
  pivot_longer(cols = `Government`:`DAH`)
```

```{r}
# table
reactable(the, filterable = TRUE)

# plot
ggplot(data = the, 
       aes(x = gdp.per.capita, y = value)) + 
  geom_point() + 
  geom_smooth(method = "gam") + 
  facet_grid(year ~ name) +
  theme_classic() +
  labs(x = "GDP per capita (2015 USD)",
       y = "Proportion of total health spending (%)") +
  ylim(0, 92)

summary(the$value)

ggplot(data = the, 
       aes(x = gdp.per.capita, y = value, color = factor(year))) + 
  geom_point() + 
  geom_smooth(method = "gam") + 
  facet_grid(~ name) +
  theme_classic() +
  scale_fill_brewer(palette = "BrBG") + #"Spectral"  
  labs(x = "GDP per capita (2015 USD)",
       y = "Proportion of total health spending (%)",
       color = "Year") +
  ylim(0, 92)

# save plot
ggsave(filename = "./report/desc_health spending.png", 
       plot = last_plot(),
       width = 18,
       height = 10,
       units = "cm",
       dpi = 600)
```

## 2.4. trend health spending: government, out-of-pocket, health aid
```{r}
# note: plot trend health spending: by components and dah

# data
## workflow
## 1. select country, year, the vars
## 2. filter non NA values
## 3. group_by year
## 4. summarize - convert to bilions 2021 USD
## 5. rename
## 6. pivot longer

trend_the <- data_explanatory %>% 
  select(country, year, gdp.per.capita, gdp, the_total:dah_total) %>%
  filter(is.na(the_total) == F, is.na(gdp.per.capita) == F) %>%
  group_by(year) %>%
  summarize(the_total = sum(the_total, na.rm = T)*1000/1000000000,
            ghes_total = sum(ghes_total, na.rm = T)*1000/1000000000,
            ppp_total = sum(ppp_total, na.rm = T)*1000/1000000000,
            oop_total = sum(oop_total, na.rm = T)*1000/1000000000,
            dah_total = sum(dah_total, na.rm = T)*1000/1000000000) %>%
  rename(`Total health spending` = the_total,
         `Government` = ghes_total,
         `Prepaid private` = ppp_total,
         `Out-of-pocket` = oop_total,
         `DAH` = dah_total) %>%
  pivot_longer(cols = `Total health spending`:`DAH`) %>%
  group_by(year) %>%
  mutate(ratio_the = round(value/max(value)*100, 1))
```

```{r}
# table
reactable(trend_the, filterable = TRUE)

# plot trend health spending: by components 
ggplot() + 
  geom_area(data = trend_the %>% filter(name != "Total health spending"), 
       aes(x = year, y = value, fill = factor(name))) +
  theme_classic() +
  labs(x = "Year",
       y = "Health spending (billions 2021 USD)",
       fill = "Spending source") +
  scale_fill_brewer(palette = "BrBG") #Set2, #Pastel1

# save plot
ggsave(filename = "./report/desc_health spending_trend.png", 
       plot = last_plot(),
       width = 18,
       height = 10,
       units = "cm",
       dpi = 600)
```

```{r}

# plot trend health spending: dah
ggplot(data = trend_the %>% filter(name == "DAH"), 
       aes(x = year, y = value)) + 
  geom_line() +
  theme_classic() +
  labs(x = "Year",
       y = "DAH (billions 2021 USD)") 

# save plot
ggsave(filename = "./report/desc_DAH_trend.png", 
       plot = last_plot(),
       width = 18,
       height = 10,
       units = "cm",
       dpi = 600)
```

## 2.5. Additional 

### DAH by country
```{r}
# dah - scaled (mean 0, sd 1)
dah <- data_explanatory %>% 
  filter(is.na(total.aid.for.health) == F) %>%
  group_by(country) %>% 
  mutate(dah = scale(total.aid.for.health, scale = T)) %>%
  select(country, year, dah, total.aid.for.health)

ggplot(data = dah, 
       aes(x = year, y =  dah)) + 
  geom_line() +
  facet_wrap(~ country) +
  theme_classic() +
  labs(x = NULL,
       y = "DAH per GDP (scaled)") +
  scale_x_continuous(guide = guide_axis(angle = 45))


ggsave(filename = "./report/desc_DAH scaled_by country.png", 
       plot = last_plot(),
       width = 40,
       height = 20,
       units = "cm",
       dpi = 600)
```

### DAH volatility by country
```{r}
# dah.vol - scaled (mean 0, sd 1)
dah.vol <- data_explanatory %>% 
  filter(is.na(total.aid.for.health.volatility) == F) %>%
  group_by(country) %>% 
  mutate(dah.vol = scale(total.aid.for.health.volatility, scale = T)) %>%
  select(country, year, dah.vol, total.aid.for.health.volatility)

ggplot(data = dah.vol, 
       aes(x = year, y =  dah.vol)) + 
  geom_line() +
  facet_wrap(~ country) +
  theme_classic() +
  labs(x = NULL,
       y = "DAH volatility (scaled)") +
  scale_x_continuous(guide = guide_axis(angle = 45))

ggsave(filename = "./report/desc_DAH volatility scaled_by country.png", 
       plot = last_plot(),
       width = 20,
       height = 10,
       units = "cm",
       dpi = 600)
  

```

### health spending per the by country
```{r}
trend_the <- data_explanatory %>% 
  select(country, year, gdp.per.capita, gdp, the_total:dah_total) %>%
  filter(is.na(the_total) == F, is.na(gdp.per.capita) == F) %>%
  group_by(year, country) %>%
  summarize(the_total = sum(the_total, na.rm = T)*1000/1000000000,
            ghes_total = sum(ghes_total, na.rm = T)*1000/1000000000,
            ppp_total = sum(ppp_total, na.rm = T)*1000/1000000000,
            oop_total = sum(oop_total, na.rm = T)*1000/1000000000,
            dah_total = sum(dah_total, na.rm = T)*1000/1000000000) %>%
  rename(`Total health spending` = the_total,
         `Government` = ghes_total,
         `Prepaid private` = ppp_total,
         `Out-of-pocket` = oop_total,
         `DAH` = dah_total) %>%
  pivot_longer(cols = `Total health spending`:`DAH`) %>%
  group_by(year, country) %>%
  mutate(ratio_the = round(value/max(value)*100, 1))
```

```{r}
# plot trend health spending: dah
ggplot(data = trend_the %>% filter(name != "Total health spending"), 
       aes(x = year, y = ratio_the, color = name)) + 
  geom_line() +
  theme_classic() +
  labs(x = NULL,
       y = "Proportion of total health spending (%)",
       color = "Spending source") +
  facet_wrap(~ country) +
  scale_x_continuous(guide = guide_axis(angle = 45))

# save plot
ggsave(filename = "./report/desc_health spending per the_trend.png", 
       plot = last_plot(),
       width = 40,
       height = 20,
       units = "cm",
       dpi = 600)
```

